The Eclipse Integrated Computational Environment

### 2017-05-30 src/content.tex
Synced additional changes from Sam. Signed-off-by: Jay Jay Billings <billingsjj@ornl.gov>

In previous work, Billings et al. interviewed simulation and modeling 
software users to compile a list of requirements for implementing and 
using these kinds of applications. In the process, they discovered 
that many of the difficulties inherent in using
high-performance modeling and simulation software fall into five
distinct categories \cite{billings_designing_2009}. These activities,
detailed in Section \ref{workflow-model}, include (1) creating input,
(2) executing jobs, (3) analyzing results, (4) managing data, and (5)
modifying code. There are many tools that address these problems
individually, but the same research found that the excess number and
specialization of these tools also contribute to the learning curve.

@@

This section concludes with an introduction of the workflow model addressed by
ICE. Section \ref{software-description} details the software
from an architectural perspective, while Section \ref{illustrative-examples}
provides a set of comprehensive examples. A
presentation of the impact is included in Section \ref{impact}, and sample 
code is provided in Section \ref{sample-code-tutorials-and-other-resources}.

@@

Individual components of workflows (i.e., workflow ``tasks'' or
``nodes'') are either encoded directly in the workflow's subclass of
Item or provided as ``Actions'' that are dynamically registered with
an ``Action Factory'' and obtained at runtime. Table 1 describes the
differences between Items, Actions, and Forms.\footnote{An upcoming 
update to the API will include the formal introduction of
IWorkflow, IWorkflowTask, and IWorkflowEngine interfaces to bring ICE's
API language closer to other systems. This must be done with care to
satisfy backwards compatibility requirements on the present API.}

@@

This design is very important. First, it means that all workflows in
ICE, regardless of their actual details and functions, can only behave in
a specific set of known and predictable ways, and this predictability 
simplifies the way the Core manages and interacts with Items. Second, by
formalizing state and error checks, it explicitly delineates which workflows 
can be executed from those that must receive additional configuration. 
Finally, it allows developers implementing Items and
Actions to specify, by contract, what is required before proceeding to the
next task, processing the workflow, or declaring a successful execution.

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{images/jaded.png}
\caption{ICE workbench in JADE for quantum computing simulations.}
\end{figure}

@@

Modeling and simulation workflows for computational scientists differ
greatly from those of experimentalists or those who primarily interact with
grid-based workflow management systems. The Eclipse ICE effort described here 
has been used to address interdisciplinary problems in 
modeling and simulation for energy science. Of particular interest are the
differences in architecture between a workflow management system focused
on modeling and simulation compared to those focused on grids. ICE's broad
applicability across many topics in energy science suggests that
there are opportunities for these systems in general. Finally, one 
interesting avenue of future exploration is coupling or integrating ICE 
with other workflow tools such as Aiida, Triquetrum, Kepler, and Pegasus,
which would make it possible to combine the best of grid and modeling and
simulation workflows to address greater challenges.

@@

The authors are grateful for the assistance and support of the following
people and institutions without whom this work would not have been
possible. This includes many people who directly contributed to the
project, either in its early days as ``NiCE'' or once it moved to
Eclipse, including: Ronald Allen, Andrew Belt, Tim Bohn, David E.
Bernholdt, Erica Grant, Mike Guidry, Forest Hull, Eric J. Lingerfelt,
Sebastien Jourdain, JiSoo Kim, Allison Koenecke, Fangzhou Lin, Greg
Lyon, Tony McCrary, John M. Hetrick III, Elizabeth Piersall, Neeti
Pokhriyal, Adrian Sanchez, Claire Saunders, Nick Stanish, Matthew Wang, and
Scott Wittenberg. The support of both the NEAMS and CASL programs is also
greatly appreciated. The authors would like to acknowledge the special
contribution of the Eclipse Foundation, the Eclipse Community, the
Eclipse Science Working Group, and our many colleagues who use and
contribute to open-source projects in the Eclipse ecosystem.

>>>>

In previous work, Billings et al. interviewed modeling and simulation subject matter experts to compile a list of requirements for implementing and 
using these kinds of applications. In the process, they discovered 
that many of the difficulties inherent in using
high-performance modeling and simulation software fall into five
distinct categories \cite{billings_designing_2009}. These activities,
detailed in Section \ref{workflow-model}, include (1) creating input,
(2) executing jobs, (3) analyzing results, (4) managing data, and (5)
modifying code. There are many tools that address these problems
individually, but the same research found that the excess number and
specialization of these tools also contribute to the learning curve.

@@

This section concludes with an introduction of the ICE workflow model. Section \ref{software-description} details the software
from an architectural perspective, while Section \ref{illustrative-examples}
provides a set of comprehensive examples. A
presentation of the impact is included in Section \ref{impact}, and sample 
code is provided in Section \ref{sample-code-tutorials-and-other-resources}.

@@

Individual components of workflows (i.e., workflow ``tasks'' or
``nodes'') are either encoded directly in the workflow's subclass of
Item or provided as ``Actions'' that are dynamically registered with
an ``Action Factory'' and obtained at runtime. Table 1 describes the
differences between Items, Actions, and Forms.\footnote{An upcoming 
update to the API will include the formal introduction of
IWorkflow, IWorkflowTask, and IWorkflowEngine interfaces to bring ICE's
API language closer to other systems. This must be done carefully to
preserve backwards compatibility with the present API.}

@@

This design is very important. First, it means that all workflows in
ICE, regardless of their actual details and functions, can only behave in
a specific set of known and predictable ways, and this predictability 
simplifies the way the Core manages and interacts with Items. Second, by
formalizing state and error checks, ICE explicitly delineates which workflows 
can be executed from those that must receive additional configuration. 
Finally, it allows developers implementing Items and
Actions to specify, by contract, what is required before proceeding to the
next task, processing the workflow, or declaring a successful execution.

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{images/jaded.png}
\caption{ICE workbench in the Jade Adiabatic Development Environment (JADE) for quantum computing simulations.}
\end{figure}

@@

Modeling and simulation workflows for computational scientists differ
greatly from those of experimentalists or those who primarily interact with
grid-based workflow management systems. The Eclipse ICE effort described here 
has been used to address interdisciplinary problems in 
modeling and simulation for energy science. Of particular interest are the
differences in architecture between a workflow management system focused
on modeling and simulation compared to those focused on grids. ICE's broad
applicability across many topics in energy science suggests that
there are opportunities for these systems in general. Finally, one 
interesting avenue of future exploration is coupling or integrating ICE 
with other workflow tools such as Aiida, Triquetrum, Kepler, and Pegasus,
which would make it possible to combine the best of grid and modeling workflows and
simulation workflows to address greater challenges.

@@

The authors are grateful for the assistance and support of the following
people and institutions without whom this work would not have been
possible. This includes many people who directly contributed to the
project, either in its early days as ``NiCE'' or once it moved to
Eclipse, including: Ronald Allen, Andrew Belt, Tim Bohn, David E.
Bernholdt, Erica Grant, Mike Guidry, Forest Hull, Eric J. Lingerfelt,
Sebastien Jourdain, JiSoo Kim, Allison Koenecke, Fangzhou Lin, Greg
Lyon, Tony McCrary, John M. Hetrick III, Elizabeth Piersall, Neeti
Pokhriyal, Adrian Sanchez, Claire Saunders, Nick Stanish, Matthew Wang, and
Scott Wittenberg. The support of both the Nucleare Energy Advanced Modeling and Simulation (NEAMS) and Consortium for Simulation of Light-water Reactors (CASL) programs is also
greatly appreciated. The authors would like to acknowledge the special
contribution of the Eclipse Foundation, the Eclipse Community, the
Eclipse Science Working Group, and our many colleagues who use and
contribute to open-source projects in the Eclipse ecosystem.
